---
title: "Geography 176A"
author: "[Yan Wang](https://vialawaila.github.io/)"
subtitle: 'Lab 03: COVID-19 Pandemic'
output:
  html_document:
    theme: journal
---

#Yan Wang
#08/17/2020
#Lab 02

```{r}
library(tidyverse)
url = 'https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv'
covid = read_csv(url)
head(covid)
```

#Question 1

```{r}
library(tidyverse)
dat <- covid %>%
  filter(state == "California") %>%
  group_by(county) %>%
  mutate(newCases = cases - lag(cases)) %>%
  ungroup() %>%
  filter(date == max(date))

most_cases = dat %>%
  slice_max(cases, n = 5) %>%
  select(county, cases)

knitr::kable(most_cases,
             caption = "Most Cases California Counties",
             col.names = c("County", "Cases"))
```


```{r}
most_new_cases = dat %>%
  slice_max(newCases, n = 5) %>%
  select(county, newCases)

knitr::kable(most_new_cases,
             caption = "Most New Cases California Counties",
             col.names = c("County", "New Cases"))
```


```{r}
library(readxl)
StatePopulationEstimates <- read_excel("~/github/geog-176A-labs/data/PopulationEstimates.xls", skip = 2) %>% 
  select(pop19 = POP_ESTIMATE_2019, fips = FIPStxt)
covid_population <- inner_join(covid, StatePopulationEstimates, by = 'fips')

most_cases_percapita <- covid_population %>% 
  filter(date == max(date)) %>% 
  filter(state == 'California') %>% 
  mutate(casesPerCapita = (cases / pop19)) %>% 
  arrange(-casesPerCapita) %>% 
  head(5)
knitr::kable(most_cases_percapita, caption = "Most Cumulative Cases Per Capita", col.names = c('Date', 'County', 'State', 'FIPS', 'Cases', 'Deaths', 'Population', 'Cases Per Capita'))

```






##Question1(10)
### (1) Describe the total number of cases
```{r}
library(tidyverse)
url = 'https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv'
covid = read_csv(url)
head(covid)
dat = covid %>%
  filter(state == "California") %>%
  group_by(county) %>%
  mutate(newCases = cases - lag(cases)) %>%
  ungroup() %>%
  filter(date == max(date))
```

```{r}
library(tidyverse)
(total_state_cases = dat %>%
    filter(date == max(date)) %>%
    group_by(county) %>%
    summarise(cases = sum(cases, na.rm = TRUE)) %>%
    ungroup() %>%
    summarise(cases = sum(cases, na.rm = TRUE)) %>%
    pull(cases))
```
##Question1(10)
### (2) Describe the total number of new cases
```{r}
library(tidyverse)
(total_state_newCases = dat %>%
    filter(date == max(date)) %>%
    group_by(county) %>%
    summarise(newCases = sum(newCases, na.rm = TRUE)) %>%
    ungroup() %>%
    summarise(newCases = sum(newCases, na.rm = TRUE)) %>%
    pull(newCases))
```
##Question1(10)
### (3) Describe the total number of safe counties 

```{r}
library(readxl)
library(tidyverse)
pop <- read_excel("../data/PopulationEstimates.xls", skip = 2)
(pop3 = pop  %>%
    filter(State == "CA") %>%
    select(pop19 = POP_ESTIMATE_2019, state = State, county = Area_Name, fips = FIPStxt) %>%
    group_by(county) %>%
    slice_max(pop19, n=1))

(dat2 = covid %>%
    filter(state == "California") %>%
    group_by(county) %>%
    mutate(newCases = cases - lag(cases)) %>%
    ungroup())
pop_dat2 = right_join(pop3, dat2, by = "fips")

(last14Days = pop_dat2 %>%
    filter(date > max(date) - 14, date < max(date)) %>%
    select(county = county.y, newCases, pop19, date) %>%
    group_by(county, pop19) %>%
    summarise(newCases = sum(newCases, na.rm = TRUE)) %>%
    ungroup() %>%
    mutate(newCases_percapita = (newCases/(pop19/100000))))
```


```{r}
library(tidyverse)
(safe_counties = last14Days %>%
    filter(newCases_percapita < 100) %>%
    pull(county))
```
**As of August 17, 2020, there are a total of 628508 cases, 6527 new cases within the state of California, and 13 counties in California are safe.**

#Question 2
```{r}
library(ggthemes)
library(zoo)
library(tidyverse)
covid %>%
  filter(state %in% c('New York', 'California', 'Louisiana', 'Florida')) %>%
  group_by(state, date) %>%
  summarize(cases = sum(cases)) %>%
  ungroup(state, date) %>%
  group_by(state) %>%
  mutate(newCases = cases - lag(cases)) %>%
  mutate(roll7 = rollmean(newCases, 7, fill = NA, align = "right")) %>%
ggplot(aes(x = date)) +
  geom_col(aes(y = newCases), col = NA, fill = "#F5B8B5") +
  geom_line(aes(y = roll7), col = "darkred", size = 1) +
  labs(title = "New Cases: States",
       x = 'Date',
       y = "Daily New Cases Count",
       caption = "Geog 176A-Lab 02",
       subtitle = "COVID-19 Data: NY-Times",
       color = "") +
  facet_wrap(~state, scales = "free_y") +
  theme(plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill = "white"),
        plot.title = element_text(size = 15, face = 'bold')) +
  theme(aspect.ratio = .5)
```


```{r}
library(tidyverse)
dat2 = covid%>%
  filter(state %in% c('New York', 'California', 'Louisiana', 'Florida')) %>%
  group_by(state, date) %>%
  summarize(cases = sum(cases)) %>%
  ungroup(state, date) %>%
  group_by(state) %>%
  mutate(NewDailyCases = cases - lag(cases)) %>%
  mutate(SevenDayAvg = rollmean(NewDailyCases, 7, fill = NA, align = "right"))
StatePopulationEstimates1 <- read_excel("~/github/geog-176A-labs/data/PopulationEstimates.xls", skip = 2) %>% 
  select(pop19 = POP_ESTIMATE_2019, state_abbr = State, state = Area_Name) %>% 
  group_by(state_abbr) %>% 
  slice_max(pop19, n = 1)
newjoineddata = inner_join(dat2, StatePopulationEstimates1, by = 'state')
percapdata <- newjoineddata %>% 
  group_by(state) %>% 
  mutate(NewCasesPerCap = NewDailyCases / pop19) %>% 
  mutate(NewSevenDayAvg = rollmean(NewCasesPerCap, 7, fill = NA, align = "right")) 
percapdata %>% 
  ggplot(aes(x = date)) +
  geom_col(aes(y = NewCasesPerCap), col = "#F5B8B5") +
  geom_line(aes(y = NewSevenDayAvg), col = "darkred", size = 1) +
  labs(title = "New Cases Per Capita: States",
       x = "Date",
       y = "Newcases",
       caption = "Geog 176A-Lab 02",
       subtitle = "COVID-19 Data: NY-Times",
       color = "") +
  facet_wrap(~state, scales = "free_y")
  theme(plot.background = element_rect(fill = "white"),
      panel.background = element_rect(fill = "white"),
      plot.title = element_text(size = 15, face = 'bold')) +
  theme(aspect.ratio = .5)

```




**Scaling by population, it data presents the directivity of the data. It becomes better for comparing the data because for some areas with absolute small population and small amount of confirmed cases (i.e. Louisiana), the ration of cases over population introduces a objective comparison between different places.It is now comparing the relative amount among different areas.**


#Question3

```{r}
library(readr)
county_centroids <- read_csv("../data/county-centroids.csv")
head(county_centroids)
```

```{r}
library(tidyverse)
library(ggplot2)
library(ggthemes)
county2 = county_centroids %>% 
  select(county = name, LON, LAT, fips) 
covid_xy = inner_join(covid, county2, by = c("county", "fips"))
xy1 <- covid_xy %>% 
  mutate(xcoord = cases * LON, ycoord = cases * LAT) %>% 
  group_by(date) %>% 
  summarize(cases = sum(cases), xcoord = sum(xcoord), ycoord = sum(ycoord)) %>% 
  mutate(longitude = xcoord / cases, latitude = ycoord / cases) %>% 
  mutate(month = format(date, "%m")) 
xy2 <- xy1 %>% 
  group_by(month) %>% 
  summarise(mocases = sum(cases))
xy3 <- inner_join(xy1, xy2, by = "month") %>%
  select(date, longitude, latitude) 
knitr::kable(xy3, caption = "COVID-19 Weighted Mean", col.names = c("Date","Longitude","Latitude"))

```


```{r}
xy4 <- xy2 %>% 
  select(month, mocases)
knitr::kable(xy4, caption = "Monthly New Cases", col.names = c("Month","New Cases"))
ggplot(data = xy1, aes(x = longitude, y = latitude)) +
  borders("state", fill = "gray", colour = "white") +
  geom_point(aes(color = month, size = cases)) +
  labs(title = "COVID-19 Weighted Mean",
       x = "Longitude",
       y = "Latitude",
       caption = "Geog 176A-Lab 02",
       subtitle = "COVID-19 Data: NY-Times",
       color = "")
```

**Weight is a relative concept. The weight of the weighted average reflects the relative importance in the overall evaluation. The weight indicates that in the evaluation process, it is the quantitative allocation of the importance degree of different aspects of the evaluated object, and the role of each evaluation factor in the overall evaluation is treated differently. In fact, an evaluation without a focus is not an objective evaluation. Weight indicates how important certain data is in a set of data, so the weighted average effect must be studied in combination with specific examples. The size of the weighted average is not only related to each data in a set but is also affected by the weight of each data. The greater the weight.The greater the effect on the average size. The reverse is smaller.**







